Theoretical study on grazing-incidence x-ray scattering of surfaces upon high-intensity laser irradiation

### 2018-09-28 paper.tex
Improved introduction (WIP).

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers after \end{frontmatter}.
%% \usepackage{lineno}

@@

\journal{Journal Name}

@@

\author[1]{M. Banjafar}
\author[2]{L. Randolph}
\author[3]{T. Kluge}
\author[3]{M. Bussmann}
\author[1]{A. Mancuso}
\author[3]{T. E. Cowan}
\author[1]{C. Fortmann-Grote}
\author[2]{C. Gutt}
\author[1]{M. Nakatsutsumi}

@@

\begin{abstract}
%% Text of abstract
  Interaction of solid materials with laser pulses of \SI{1e16}{\watt\per\cm}
  produces warm dense matter under controlled laboratory conditions.
  Multilayer targets of alternating high and low Z material layers of a few
  nanometer thickness open the possibility
  to study the laser--matter interaction dynamics as a volumetric effect by
  observing the multilayer Bragg peak in a x--ray reflectometry experiment.
  Here, we present particle--in--cell simulations of
  the laser and multilayer target interaction and subsequent calculations of
  the reflectivity as a
  function of reflection angle and various time delays between the optical pump
  and the x--ray probe pulse delivered by an x--ray free--electron laser.
  Using the same methodology, we also calculate the diffuse x--ray scattering
  signal.
  We demonstrate that the combined measurement of reflectivity and
  diffuse scattering provides valuable information about the evolution of
  electron density gradients, temperature and ionization as a function of target
  depth and pump--probe time delay and Bragg peak position.
\end{abstract}

\begin{keyword}
Science \sep Publication \sep Complicated
%% keywords here, in the form: keyword \sep keyword

%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)

\end{keyword}

\end{frontmatter}

%%
%% Start line numbering here if you want
%%
%%\linenumbers
\begin{verbatim}
Carsten:
* Abstract to be iterated when rest of the text comes into shape.}
* What figures will go in the paper?
Thomas:
  * Experimental description should focus on giving the physical
  motivation of why we want to do surface diffraction at all
  * Simulation part should focus on showing
    (1) the qualitative plasma physics we expect
    (2) the differences between codes with respect to
      (a) different physics included (e.g. comparing with and without
      ionization, with and withou collisions etc)
      (b) different models, e.g.
        i) TF vs. Direct impact ionization
        ii) large collision frequency vs. low collision frequency
        iii) comparing a constant Coulomb logarithm vs. a T-dependent one
        iV) hydro vs. PIC
      (c) different model implementations (e.g. TF and ADK are different
      in PICLS and PIConGPU)
    (3) lay the foundation in showing what plasma physics is important and
    what the uncertainties in the numerical modeling are, so that in the
    scattering part one can focus on showing how to measure the physics
    and validate the codes
  * The scattering part has to derive much more stringent the connection
  between the measured quantities (e.g. beta, change of max(Q) position,
  fourier transform/rings) and plasma physical quantities (e.g. what
  exactly are slow/fast dynamics and how do they imprint on the speckle
  contrast? surface ripple correlation changes - where do they come from
  (should be explained in the sim-part) and how do the show up in the
  signal, what exactly expands and how/why does it imprint on the signal,
  what is the plasma physics measured by the ring-analysis?...)
\end{verbatim}
\begin{verbatim}
TODOS:
  * Add authors, keywords
  * Abstract
  * Find out how many pages we have for the RPHDM proceedings? Depending on this
  info, structure the paper, suggestions below
  * Introduction:
    -Motivation
    - Intensity regime -> WDM, some words about prepulse in high-power laser
    experiments
  * Schematic experiment setup (use figure from talk)
  * Methods:
    - PIC: which code(s), only short, give references. briefly discuss
  collisional ionization and collisions.
    - reflectometry calculations from PIC results
  * Results:
    - electron density plots, maybe video for supplemental material (only Ta-Al,
    or also Ta-Si?)
    - electron density profiles, difference betwen codes (do we want to show
    here, is it checked?)
    - reflection curves, diffuse scattering (?)
  * Discussion
  * Summary, Outlook
\end{verbatim}

@@

%% motoaki
(1) High-intensity laser-matter (solid) interaction opens broad applications such as ... bra bra ...
(2) Laser-matter coupling is largely dominated by instantaneous electron density

%%

The warm dense matter regime is a state of matter in which the materials have been exposed to the high pressure of 1 Mbar and have the properties of both plasma and solid state. In such a condition the temperature of the electrons varies from a few eV to 100 eV and the density is typically from solids up to highly compressed. These unique physical properties of WDM and its importance and practical applications in physics of planetary formation like the interior layers structure, convection, magnetic field, and charge transfer inside the gas giants and the core of Earth-like planets [], inertial confinement fusion where the material is encountered during the implosion phase [], plasma optics [], and material science [] have made it an exciting topic to study. However, this unique position of WDM in density and temperature space makes this area of physics very challenging to investigate. Actually, this state of matter can be produced in the laboratory by irradiating a very thin foil with a few tens of nm thickness by femtosecond optical laser pulses with $10^{15}-10^{16}$ $W/cm^2$ intensity []. But, for usual optical diagnostic techniques, WDM acts like an opaque because of the high density of free electrons. Thus, in order to obtain physical information like density profile and temperature from inside the WDM other techniques such and proton and X-ray radiography have been commonly using [].

Some novel techniques in laser--plasma interactions such and plasma mirrors[], in which a femtosecond laser pulse is efficiently reflected by the dense plasma with higher temporal contrast and tighter focusing, and using pre-pulses [], which often precede the main high-intensity laser pulse, are categorized in the WDM regime due to the up to $10^{16}$ $W/cm^2$ intensity of the laser pulses that is commonly used in these technique. In these cases, having a good knowledge of the surface structure such as nano-scale roughness and surface dynamics like plasma hydrodynamic expansion is very important to optimize the experiments. Here, X-ray free electron laser (XFEL) has the capability to provide high temporal (femtosecond) and spacial (nanometer) resolution that is crucial to get some information from the surface of hot dense plasma. It also provides a very good vision of electron density gradient and the opacity [] of warm dense plasma. Some X-ray probing techniques such as X-ray reflectivity (XRR) and grazing incidence X-ray diffraction (GIXD) [] are particularly sensitive to the sample electron distribution in the plane and in the direction perpendicular to the substrate surface, by which we are able to simply obtain the depth-resolved structure characterization by altering the incidence angle.

In this work, we present Particle--in--Cell simulation and X--ray reflectometry study of interaction of a $10^{16}$ $W/cm^2$ intensity optical laser pulse with a multilayer target based on an experiment with the same geometry.

@@

\section{Modeling}
\label{S:2}
\subsection{Particle--in--Cell simulation}
A schematic of 2D PIC simulation setup has been shown in (figure). The simulation was done by PIConGPU code [reference] on $4096 \times 4096$ cells that the dimension of each cells is 2.5 {\AA} in both direction and the time step is ${\Delta}t=5.89 \times 10^{-4}$ fs. Thus, the dimension of simulation box is nearly $1 \times 1$ $\mu m$ and the total time of simulation is 527 fs. An optical laser pulse with plane wave profile, intensity $I = 10^{16}$ $W/{cm}^2$ corresponds to $a_{0} = 0.068$ for 800 $nm$ laser wavelength, and the pulse duration $\tau_{FWHM}=50$ fs, and focal spot size of 5 $\mu m$  comes alongside the $y$ axis and incidents normally on the front of multilayer (ML) target, which is a 10 times reps of Tantalum, $n_{Ta} = 5.55 \times 10^{22}$ $cm^{-3}$, and Aluminum ,$n_{Al} = 6.022 \times 10^{22}$ $cm^{-3}$, double-layers. The thickness of each layer is 5 $nm$ and regarding the 20 $nm$ Silicon substrate, the total thickness of the target is 120 $nm$. Three kinds of ionization model, barrier suppression ionization (BSI) [reference], Thomas-Fermi (TF) [reference], and the tunnel ionization (ADK) [reference] have been used in our simulation, while it dose not include the collision process. In fact, quantitative modeling of WDM is a critical challenge for kinetic method such as particle--in--cell and radiation--hydrodynamics mainly due to (a) difficulties in describing the ionization dynamics; (b) uncertainty about electron-ion collision frequency which governs absorption, heat transfer and plasma conductivity; and (c) violation of the heat conduction models at the steeped temperature gradient present in laser excited plasma []. However,

@@

\section{Acknowledgment}
text
\newline
\newline
\newline
\newline
\newline

To add the equations, tables, plots, and bullet points use these items:

\begin{equation}
\label{eq:emc}
e = mc^2
\end{equation}

\begin{figure}[h]
\centering\includegraphics[width=0.4\linewidth]{placeholder}
\caption{Figure caption}
\end{figure}

\begin{table}[h]
\centering
\begin{tabular}{l l l}
\hline
\textbf{Treatments} & \textbf{Response 1} & \textbf{Response 2}\\
\hline
Treatment 1 & 0.0003262 & 0.562 \\
Treatment 2 & 0.0015681 & 0.910 \\
Treatment 3 & 0.0009271 & 0.296 \\
\hline
\end{tabular}
\caption{Table caption}
\end{table}

\begin{itemize}
\item Bullet point one
\item Bullet point two
\end{itemize}

\begin{enumerate}
\item Numbered list item one
\item Numbered list item two
\end{enumerate}

%% The Appendices part is started with the command \appendix;
%% appendix sections are then done as normal sections
%% \appendix

%% \section{}
%% \label{}

%% References
%%
%% Following citation commands can be used in the body text:
%% Usage of \cite is as follows:
%%   \cite{key}          ==>>  [#]
%%   \cite[chap. 2]{key} ==>>  [#, chap. 2]
%%   \citet{key}         ==>>  Author [#]

%% References with bibTeX database:

\bibliographystyle{model1-num-names}
\bibliography{sample.bib}

%% Authors are advised to submit their bibtex database files. They are
%% requested to list a bibtex style file in the manuscript if they do
%% not want to use model1-num-names.bst.

%% References without bibTeX database:

% \begin{thebibliography}{00}

%% \bibitem must have the following form:
%%   \bibitem{key}...
%%

% \bibitem{}

% \end{thebibliography}

>>>>

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers after \end{frontmatter}.
\usepackage{lineno}
\usepackage[]{hyperref}

@@

\DeclareSIUnit{\Wsqcm}{\watt\per\square\centi\meter}
\journal{Journal Name}

@@

\author[1]{M. Banjafar}
\author[2]{L. Randolph}
\author[3]{T. Kluge}
\author[3]{M. Bussmann}
\author[1]{A .P. Mancuso}
\author[3]{T. E. Cowan}
\author[1]{C. Fortmann-Grote}
\author[2]{C. Gutt}
\author[1]{M. Nakatsutsumi}

@@

\begin{abstract}
%% Text of abstract
  Interaction of solid materials with laser pulses
  produces warm dense matter under controlled laboratory conditions.
  Multilayer targets of alternating high and low Z material layers of a few
  nanometer thickness open the possibility
  to study the laser--matter interaction dynamics as a volumetric effect by
  observing the multilayer Bragg peak in a x--ray reflectometry experiment.
  Here, we present particle--in--cell simulations of
  the laser and multilayer target interaction and subsequent calculations of
  the reflectivity as a
  function of reflection angle and various time delays between the optical pump
  and the x--ray probe pulse delivered by an x--ray free--electron laser.
  Using the same methodology, we also calculate the diffuse x--ray scattering
  signal.
  We demonstrate that the combined measurement of reflectivity and
  diffuse scattering provides valuable information about the evolution of
  electron density gradients, temperature and ionization as a function of target
  depth and pump--probe time delay and Bragg peak position.
\end{abstract}

\begin{keyword}
Short pulse laser--matter interaction \sep particle--in--cell simulations \sep x--ray
diagnostics \sep x--ray free--electron laser
%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)
\end{keyword}
\end{frontmatter}

%%
%% Start line numbering here if you want
%%
\linenumbers
\begin{verbatim}
Carsten:
* Abstract to be iterated when rest of the text comes into shape.}
* What figures will go in the paper?
Thomas:
  * Experimental description should focus on giving the physical
  motivation of why we want to do surface diffraction at all
  * Simulation part should focus on showing
    (1) the qualitative plasma physics we expect
    (2) the differences between codes with respect to
      (a) different physics included (e.g. comparing with and without
      ionization, with and withou collisions etc)
      (b) different models, e.g.
        i) TF vs. Direct impact ionization
        ii) large collision frequency vs. low collision frequency
        iii) comparing a constant Coulomb logarithm vs. a T-dependent one
        iV) hydro vs. PIC
      (c) different model implementations (e.g. TF and ADK are different
      in PICLS and PIConGPU)
    (3) lay the foundation in showing what plasma physics is important and
    what the uncertainties in the numerical modeling are, so that in the
    scattering part one can focus on showing how to measure the physics
    and validate the codes
  * The scattering part has to derive much more stringent the connection
  between the measured quantities (e.g. beta, change of max(Q) position,
  fourier transform/rings) and plasma physical quantities (e.g. what
  exactly are slow/fast dynamics and how do they imprint on the speckle
  contrast? surface ripple correlation changes - where do they come from
  (should be explained in the sim-part) and how do the show up in the
  signal, what exactly expands and how/why does it imprint on the signal,
  what is the plasma physics measured by the ring-analysis?...)
\end{verbatim}
\begin{verbatim}
TODOS:
  * Abstract
  info, structure the paper, suggestions below
  * Introduction:
    -Motivation
    - Intensity regime -> WDM, some words about prepulse in high-power laser
    experiments
  * Schematic experiment setup (use figure from talk)
  * Methods:
    - PIC: which code(s), only short, give references. briefly discuss
  collisional ionization and collisions.
    - reflectometry calculations from PIC results
  * Results:
    - electron density plots, maybe video for supplemental material (only Ta-Al,
    or also Ta-Si?)
    - electron density profiles, difference betwen codes (do we want to show
    here, is it checked?)
    - reflection curves, diffuse scattering (?)
  * Discussion
  * Summary, Outlook
\end{verbatim}

@@

High--intensity laser--matter interaction with solid density targets opens
broad applications such as high--harmonic generation \cite{}, laboratory
astrophysics \cite{}, materials research \cite{}, laser driven fusion
\cite{}, particle acceleration \cite{}, and optical applications such as plasma
mirrors \cite{}.
Laser--matter coupling is largely dominated by the instantaneous electron density.
Substantial ionization due to \ldots can already be observed at
non--relativistic laser intensities (\SI{\ll
e18}{\Wsqcm}). A femtosecond optical laser pulse at moderate intensities
(\SIrange{e14}{e16}{\Wsqcm}) impinging on a thin
metal foil with a few tens of \si{\nm} thickness produces pressures in excess of
\SI{1}{\mega\bar}. Under such conditions, warm dense matter (WDM) forms
with electron densities close to solid density and temperatures of a few
thousand to a few ten thousand \si{\K}. This is also the mechanism of a plasma
mirror, where a pre--pulse of  \SI{\approx e16}{\Wsqcm} makes the target surface
highly reflective and thereby improves the temporal contrast and focussing of
the main pulse \marginpar{Is this correct?}.
It is thus of strong interest to understand
the physics of WDM and in particular the spatial structure on a \si{\nm} scale
and the electron dynamics in the \si{\fs} regime
because these in turn determine the interaction with the peak
part of the relativistic pulse.
Beyond this technological motivation, there is also fundamental interest in WDM
due to its abundance in many astrophysical objects \cite{Nettelmann2008a}
such as planetary
interiors (including the earth's), and stellar atmospheres. The regime of WDM
is also crossed by the trajectory of inertial confinement fusion targets \cite{Lindl2004}, adding
again to the technologically driven interest in this subject.

WDM is a particularly challenging state of matter both from the experimental and from the
theoretical perspective. Being a highly transient state, extremely high temporal
and spatial resolution is required to probe the target's structure and dynamics
using e.g. pulsed x--ray sources. Recently, laser--plasma driven x--ray sources
and free--electron lasers have become the state--of--the--art sources for this
research \cite{Lee2003, Lee2003a}.
X--ray reflectometry (XRR) and grazing incidence x--ray diffraction (GIXD) \cite{TBC} are particularly
sensitive to the sample electron distribution in the plane and in the direction
perpendicular to the target surface.
<+Discuss some difficulties, complex
dynamics+>.

The interpretation and analysis of experimental data
rely on theoretical models
and numerical simulations which have their own difficulties.
This is mainly due to WDM being located in a peculiar region of the density--temperature phase
diagram with temperatures comparable to both the electrons' Fermi energy and the
potential energy. These conditions give rise to partial quantum degeneracy and
moderate coupling. On top, the electron and ion subsystems often have quite distinct
temperatures and densities, requiring different theoretical approaches to treat
either one.

In
Particle--in--cell simulations \cite{} have become the state--of--the--art
numerical technique for hot, strongly ionized plasmas. In the WDM regime, where
collisons and atomic physis become important, the PIC method require substantial
improvements to account for this physics. This is an active area of research.
In this work, we present Particle--in--Cell simulation and x--ray reflectometry
study of interaction of a $10^{16}$ $W/cm^2$ intensity optical laser pulse with
a multilayer target based on an experiment with the same geometry.

@@

\section{Modeling}
\label{S:2}
\subsection{Particle--in--Cell simulation}
A schematic of 2D PIC simulation setup has been shown in (figure). The
simulation was done by PIConGPU code [reference] on $4096 \times 4096$ cells
that the dimension of each cells is 2.5 {\AA} in both direction and the time
step is ${\Delta}t=5.89 \times 10^{-4}$ fs. Thus, the dimension of simulation
box is nearly $1 \times 1$ $\mu m$ and the total time of simulation is 527 fs.
An optical laser pulse with plane wave profile, intensity $I = 10^{16}$
$W/{cm}^2$ corresponds to $a_{0} = 0.068$ for 800 $nm$ laser wavelength, and the
pulse duration $\tau_{FWHM}=50$ fs, and focal spot size of 5 $\mu m$  comes
alongside the $y$ axis and incidents normally on the front of multilayer (ML)
target, which is a 10 times reps of Tantalum, $n_{Ta} = 5.55 \times 10^{22}$
$cm^{-3}$, and Aluminum ,$n_{Al} = 6.022 \times 10^{22}$ $cm^{-3}$,
double-layers. The thickness of each layer is 5 $nm$ and regarding the 20 $nm$
Silicon substrate, the total thickness of the target is 120 $nm$. Three kinds of
ionization model, barrier suppression ionization (BSI) [reference], Thomas-Fermi
(TF) [reference], and the tunnel ionization (ADK) [reference] have been used in
our simulation, while it dose not include the collision process. In fact,
quantitative modeling of WDM is a critical challenge for kinetic method such as
particle--in--cell and radiation--hydrodynamics mainly due to (a) difficulties
in describing the ionization dynamics; (b) uncertainty about electron-ion
collision frequency which governs absorption, heat transfer and plasma
conductivity; and (c) violation of the heat conduction models at the steeped
temperature gradient present in laser excited plasma \cite{TBC}. However,

@@

\section{Acknowledgment}
CFG acknowledges support from the European Cluster of Advanced Laser Light
Sources (EUCALL) project which has received funding from the European Union’s
Horizon 2020 research and innovation programme under grant agreement No 654220.

%\begin{equation}
%\label{eq:emc}
%e = mc^2
%\end{equation}

%\begin{figure}[h]
%\centering\includegraphics[width=0.4\linewidth]{placeholder}
%\caption{Figure caption}
%\end{figure}

%\begin{table}[h]
%\centering
%\begin{tabular}{l l l}
%\hline
%\textbf{Treatments} & \textbf{Response 1} & \textbf{Response 2}\\
%\hline
%Treatment 1 & 0.0003262 & 0.562 \\
%Treatment 2 & 0.0015681 & 0.910 \\
%Treatment 3 & 0.0009271 & 0.296 \\
%\hline
%\end{tabular}
%\caption{Table caption}
%\end{table}
%% References
%%
%% Following citation commands can be used in the body text:
%% Usage of \cite is as follows:
%%   \cite{key}          ==>>  [#]
%%   \cite[chap. 2]{key} ==>>  [#, chap. 2]
%%   \citet{key}         ==>>  Author [#]

%% References with bibTeX database:
\bibliographystyle{model1-num-names}
\bibliography{references.bib, library.bib}
